\section{Question 6 \texorpdfstring{$\checkmark$}{\text{✓}}}
\horz
Let $X$ be a finite-dimensional normed linear space and $E\subset X$. Show that $E$ is compact if and only if $E$ is closed and bounded subset of $X.$ 
\horz

\begin{proof}[Solution]
    Let $X$ be $n$ dimensional normed linear space with the norm $\norm{}_{1}$. We can also give an inner product structure $X$ in the following way: Fix a basis $\left\{ v_{i} : i\in \left\{ 1,2,\ldots , n \right\} \right\}$ of $X$ and we define an inner product:
    \begin{equation}
	\ip{\sum_{i=1}^{n} x_{i}v_{i}, \sum_{i=1}^{n} y_{i}v_{i}} = \sum_{i=1}^{n} x_{i}\overline{y_{i}}	
	\label{eqn:q6}
    \end{equation}
    Since every finite dimensional inner product space is isometrically isomorphic to $\C^{n}$ with the $2$-norm, we have that $X$ with the norm induced by the inner product defined in Equation \ref{eqn:q6} is isometrically isomorphic with $\C^{n}$ with the $2$-norm. Let's call this norm induced by the inner product by $\norm{}_{2}$. Let $T: \left( X, \norm{}_{2} \right) \to \left( \C^{n} , \norm{}_{2} \right)$ be an isometric isomorphism.

    We claim that in $\left( X, \norm{}_{2} \right)$, $E$ is compact iff $E$ is closed and bounded. Since in every metric space, we have that compact subsets are closed and bounded, we need to check only one direction. To do so, let $E$ be closed and bounded subset of $X$. Then $T(E)$ is closed and bounded because $T$ is an isometric isomorphism (and hence it is an homeomorphism and preserves length). Since $T(E)$ is a closed and bounded subset of $\C ^{n}$, we have that $T\left( E \right)$ is compact. Since $T$ is an homeomorphism, $E=T^{-1} \left( T\left( E \right) \right)$ is compact. This completes the proof of the claim.

    Now, every norm on finite dimensional vector space is equivalent, so, the topology generated by $1$-norm is the same as the topology generated by the $2$-norm. So, we have that following:
    \begin{align*}
	E \text{ is compact in } \left( X, \norm{}_{1} \right) & \Leftrightarrow E \text{ is compact in } \left( X, \norm{}_{2} \right) & \text{topologies are the same!} \\
    & \Leftrightarrow E \text{ is closed and bounded in } \left( X, \norm{}_{2} \right) & \text{by the above claim} \\
& \stackrel{(\star )}{\Leftrightarrow} E \text{ is closed and bounded in } \left( X, \norm{}_{1} \right) & \text{topologies are the same!} \\
    \end{align*}
    Observe that in $\left( \star \right)$, closed is a topological property but not boundedness. But boundness is due to definition of equivalence of norms. This completes the proof.
\end{proof}
